QPLIB: A Library of Quadratic Programming Instances

### 2018-08-18 QPLIB-1.0.tex
update todos
Section 1 typographical suggestions

Quadratic Programming (QP) problems---mathematical optimization problems for which the objective function \cite{wiki:qp}, the constraints \cite{wiki:qcqp}, or both are polynomial function of the variables of degree two---include a notably diverse set of different instances. This is not surprising, given the vast scope of practical applications of such problems, and of solution methods designed to solve them \cite{GoulToin00a}. Depending on the specifics of the formulation, solving a QP may require primarily combinatorial techniques, ideas rooted in nonlinear optimization principles, or a mix of the two. In this sense, QP is arguably one of the classes of problems where collaboration between the communities interested in combinatorial and in nonlinear optimization is most needed, and potentially fruitful.

However, this diversity also implies that QP means very different things to different researchers. This is illustrated by the fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case. Therefore, it is perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been no single library devoted to all different kinds of instances of QP.
%\todo{AG: many commas in this sentence SV: I removed one.}\ 
While several specialized libraries devoted to particular cases of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches. To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted. This work constitutes a first step in this direction.

In this paper, we report the steps that have been done to collect what we consider to be a quality library of QP instances, filtering a much larger set of currently available (or specifically provided) instances in order to end up with a manageable set that still contains a meaningful sample of possible QP types. A particularly thorny issue in this process was how to select instances that are ``interesting''. Our choice has been to take this to mean ``challenging for a significant set of solution methods''. Our filtering process has then been in part based on the idea that, if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library. Conversely, if very few (maybe one) of the solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance should be deemed ``interesting''. Putting this approach into practice requires a nontrivial number of technical steps and decisions that are detailed in the paper. We hope that our work can provide useful guidelines for other researchers interested in the constructions of benchmarks for mathematical optimization problems.

A consequence of our focus is that this paper is \emph{not} concerned with the performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them. The only reason that solvers are used (and, therefore, described) in this context is to ensure that the instances of the library are nontrivial, at least for a significant fraction of the current solution methods. Providing guidance about which solvers are most suited to some specific class of QPs is entirely outside the scope of our work.

>>>>

Quadratic Programming (QP) problems---mathematical optimization problems for which the objective function \cite{wiki:qp}, the constraints \cite{wiki:qcqp}, or both are polynomial function of the variables of degree two---include a notably diverse set of different instances. This is not surprising, given the vast scope of practical applications of such problems, and of solution methods designed to solve them \cite{GoulToin00a}. Depending on the formulation specifics, solving a QP may require primarily combinatorial techniques, ideas rooted in nonlinear optimization principles, or a mix of the two. In this sense, QP is a class of problems where collaboration between the communities interested in combinatorial and in nonlinear optimization is potentially fruitful.

However, this diversity also implies that QP means very different things to different researchers. This is illustrated by the fact that the class of problems that we simply refer to here as ``QP'' might more accurately be called Mixed-Integer Quadratically-Constrained Quadratic Programming (MIQCQP) in the most general case. Therefore, it is perhaps not surprising that, unlike for ``simpler'' problems classes like Mixed-Integer Linear Programming \cite{Koch2011},  there has been no single library devoted to all different kinds of instances of QP.
While several specialized libraries devoted to particular classes of QP are available, each of them is either focused on a particular application (a specific problem that can be modeled as a QP), or on QPs with specific structural properties that make them suitable for solution by some given class of algorithmic approaches. To the best of our knowledge, collecting a set of QP instances that is at the same time not overwhelmingly onerous but sufficiently challenging for the many different interested communities has not been attempted. This work constitutes a first step in this direction.

This paper reports our steps towards collecting what we consider to be a quality QP instance library, filtering a much larger set of currently available (or specifically provided) instances and proposing a manageable set that still contains a meaningful sample of possible QP types. A particularly thorny issue in this process was how to select instances that are ``interesting''. Our choice has been to take this to mean ``challenging for a significant set of solution methods''. Our filtering process has then been in part based on the idea that, if a significant fraction of the solvers that can solve a QP instance do so in a ``short'' time, then the instance is not challenging enough to be included in the library. Conversely, if very few (maybe one) of the solvers can solve it very efficiently by exploiting some specific structure, but most other approaches cannot, then the instance should be deemed ``interesting''. Putting this approach into practice requires a nontrivial number of technical steps and decisions that are detailed in the paper. We hope that our work can provide useful guidelines for other researchers interested in constructing benchmarks for mathematical optimization problems.

A consequence of our focus is that this paper is \emph{not} concerned with the performance of the very diverse available set of QP solvers; we will \emph{not} report any data comparing them. The only reason that solvers are used (and, therefore, described) in this context is to ensure that the library instances are nontrivial, at least for a significant fraction of the current solution methods. Providing guidance about which solvers are most suited to some specific class of QPs is entirely outside the scope of our work.

### 2018-08-18 QPLIB-3.tex
fix commented text
Section 1 typographical suggestions

%\textcolor{red}{
%One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the %library in the very same form as they have been provided to us by the original contributors. Section \ref{ssec:reform} is %crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class %to another. Tailoring the structure of a problem to a solver is, however, a bias that we did not want to add.
%}

We now report some graphs that help in illustrating the main features
of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
of variables (horizontal axis) versus the number of constrains
(vertical axis), both in logarithmic scale. Continuous instances are
marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
that the library contains a quite diverse set of instances in
terms of number of variables and constraints. {The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

>>>>

%One of the nontrivial choices in our library is that we made no effort to reformulate the instances, and inserted them in the library in the very same form as they have been provided to us by the original contributors. Section \ref{ssec:reform} is crucial in justifying this choice, as it shows that there are several degrees of freedom to move the instances from one class to another. Tailoring the structure of a problem to a solver is, however, a bias that we did not want to add.

We now report some graphs that help in illustrating the main features
of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
of variables (horizontal axis) versus the number of constraints
(vertical axis), both in logarithmic scale. Continuous instances are
marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
that the library contains a quite diverse set of instances in
terms of number of variables and constraints. {The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

### 2018-08-18 QPLIB.tex
Section 1 typographical suggestions

\begin{abstract}
 This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where the objective function, the constrains, or both are quadratic.
 QP is a very diverse class of problems, comprising sub-classes of problems ranging from
 trivial to undecidable. This diversity is reflected in the variety of solution methods for QP, ranging from
 entirely combinatorial ones to completely continuous ones, including many for
 which both aspects are fundamental. Selecting a set of instances of QP that is
 at the same time not overwhelmingly onerous but sufficiently challenging for the
 many different interested communities is therefore important. We propose a simple
 taxonomy for QP instances that leads to a systematic problem selection mechanism.
 We then briefly survey the field of QP, giving an overview of theory, methods and
 solvers. Finally, we describe how the library was put together, and detail its final
 contents.
\end{abstract}

>>>>

\begin{abstract}
 This paper describes a new instance library for Quadratic Programming (QP), i.e., the family of continuous and (mixed)-integer optimization problems where the objective function and/or the constraints are quadratic.
 QP is a very diverse class of problems, comprising sub-classes ranging from
 trivial to undecidable. This diversity is reflected in the variety of QP solution methods, ranging from
 entirely combinatorial approaches to completely continuous algorithms, including many methods for
 which both aspects are fundamental. Selecting a set of instances of QP that is
 at the same time not overwhelmingly onerous but sufficiently challenging for the
 different, interested communities is therefore important. We propose a simple
 taxonomy for QP instances leading to a systematic problem selection mechanism.
 We then briefly survey the field of QP, giving an overview of theory, methods and
 solvers. Finally, we describe how the library was put together, and detail its final
 contents.
\end{abstract}
